---
title: "GSS challange"
author: "Sisilia Martinez"
date: "6/21/2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

 The bar chat is showing us who is in favor and who opposes capital punishment. It shows the response between white, black, and other races. Looking at the bar chart you can see that white people favor capital punishment more than black people. You can also see that black people have a higher number that opposes capital punishment than other races and white people. Other races have a higher number of people who favor rather than oppose. Looking at the percentage graph you can see that the white race and other race have a higher amount of people who favor capital punishment and that the black race has more who oppose it. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

In the numerical results you can see the white race has 67% in favor and 32% opposing which is 35% difference between the two. You can also see that for the black race they have 46% favoring and 53% opposing which is a 7% difference between the two. Lastly you can see other races have 56% favoring and 44% opposing which is a 12% difference between the two. This helps proves our point that both the white race and other have a higher amount of people who favor capital punishment and that the black race has a higher amount of people who oppose capital punishment.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfy the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

The Chi-Square test does not tells us there is not a difference based off the p-value given here. So we will have to run the fisher-XY test to see if there is a difference.

```{r}
fisherXY(mydata$cappun,mydata$race,printAll = TRUE)
```
 
 Looking at the fisherXY test it tells us that when comparing all three races to one another there is a difference between all of them based on the p-value shown on the test results.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->

We can conclude that in this data set the white race and other race had a higher amount of people who favor capital punishment and that the black race had the higher amount of people who oppose capital punishment. We were able to look at the bar charts that show us the differences and the numerical results confirmed our inferences by telling us that the white race had 35% more people who favored capital punishment, the other race had 12% more people who favored capital punishment, and that the black race had 7% more people who opposed capital punishment. 
